---
title: "STAT 331 Project Group I"
format: html
editor: visual
embed-resources: true
bibliography: references.bib
---

## Datasets

Using the two datasets 'income_per_person_long_series' and 'internet users' we plan to find that the higher a persons income is the more access they can have to technology.

## Cleaning the data

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
library(tidyverse)
```


```{r}
income_original <- read_csv("income_per_person_long_series.csv")
internet_original <- read_csv("internet_users.csv")
```
## Income Cleaning
```{r}
# Function to convert 'k' values (e.g., "1.1k") to numeric
convert_k_to_num <- function(x) {
  x <- as.character(x)
  ifelse(
    str_detect(x, "k"),
    as.numeric(str_remove(x, "k")) * 1000,
    as.numeric(x)
  )
}

# Apply conversion to all columns except 'country'
income_clean <- income_original |>
  mutate(across(-country, convert_k_to_num)) |>
  pivot_longer(-country, names_to = "year", values_to = "income") |>
  mutate(year = as.numeric(year))
```

## Internet Cleaning
```{r}
internet_clean <- internet_original |>
  mutate(across(-country, convert_k_to_num)) |>
  pivot_longer(-country, names_to = "year", values_to = "internet_users") |>
  mutate(year = as.numeric(year))
```

Join the Datasets
```{r}
combined_data <- income_clean |>
  inner_join(internet_clean, by = c("country", "year"))

```

## Written Component
We are examining whether there is a relationship between a country’s GDP per capita and the percentage of its population using the internet (@Gapminder) and (@World_Bank_Open_Data). Specifically, we hypothesize that countries with higher GDP per capita will also have higher internet usage. We thought that wealthier countries are more likely to have the technological accessibility needed to support widespread internet. We did find outside evidence to support this claim, (@Author_2014) found that income is a strong predictor of internet use across countries. One thing we had to do to prepare the data for analysis involved converting GDP per capita values into full numeric values. For example, income values like "31.7k" were converted to "31700" to allow for numerical analysis.